Logsums

### 2019-07-17 03_logsums.Rmd
Readjust location of logsum calculator script

```{r setup}
library(tidyverse)
library(sf)
library(spdep)
library(leaflet)
```

@@

Now we can calculate everything.
```{r distances}
# distances from all tract centroids to all parks
distances <- st_distance(pop_weighted_centroids, open_spaces, byid = TRUE) %>%
  units::set_units(miles) %>%
  units::drop_units()

# assert that a tract must be at least 1/10 mile from a park
distances <- pmax(distances, 0.1)
```

```{r write}
write_rds(distances, "data/distances.rds")
write_rds(W, "data/W.rds")
```

## Logsums
Here's the function we use to calculate the logsums
```{r}
#' Calculate destination choice logsums from a distance matrix
#'
#' @param d An $n\times p$ matrix with the distance from all tracts to
#'   all parks
#' @param sizes A p-length vector of park sizes
#' @param tweets A p-length vector of tweets at parks
#' @param betas A vector containing the size, distance, and tweet coefficients.
#'   If we submit two variables the tweets are ignored.
#'
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a
#'
calculate_park_logsums <- function(d, sizes, tweets = NULL,
                                   betas = c(-0.4,0.2,0.001)){

  # A is n x p
  a <- betas[1] * d

  # B is p x 1
  b <- betas[2] * sizes

  if(!is.null(tweets))
    b <- b + betas[3] * tweets

  # calculate observed utility by adding the weighted park-level attributes
  # to the columns of the matrix
  # V is n x p, with b added by-column to each element in a
  V <- sweep(a, 2, b, `+`)

  # log-sum of exponentiated utility, Output is n-length vector
  log(rowSums(exp(V)))

}
```

>>>>

```{r setup}
knitr::opts_chunk$set(cache = TRUE)
library(tidyverse)
library(sf)
library(spdep)
library(leaflet)
```

@@

Now we can calculate the distance from every park to every tract population
centroid, and write the distances and neighbors weights matrix to a file.
```{r distances}
# distances from all tract centroids to all parks
distances <- st_distance(pop_weighted_centroids, open_spaces, byid = TRUE) %>%
  units::set_units(miles) %>%
  units::drop_units()

# assert that a tract must be at least 1/10 mile from a park
distances <- pmax(distances, 0.1)

write_rds(distances, "data/distances.rds")
write_rds(W, "data/W.rds")
```

## Logsums

Here's the function we use to calculate the logsums. 
```{r calculate_park_logsums}
source("R/logsums.R")
print(calculate_park_logsums)
```

We use this function with asserted beta values of -1.98 and 0.3 to calculate the
accessibility logsum.
```{r logsum_vars}
logsum_vars <- tracts %>%
  transmute(
    geoid, 
    # Log-sum based utilities, without and with Twitter data
    access_ls = calculate_park_logsums(
      log(distances) * -1.98, log(open_spaces$acres) * 0.3
    ),
    # Normalize the logsum onto the Z-scale
    access_ls_n = (access_ls - mean(access_ls) ) / sd(access_ls)
  ) %>%
  st_set_geometry(NULL)
```

We are also interested in looking at what happens when the tracts are within a 
10-minute walk of a park, the threshold established by the Trust for Public Lands.
```{r nyc10mw}
access_10mw <- st_read("data/nyc_10mw.geojson") %>%
  summarise(d = 1)

inside_access <- pop_weighted_centroids %>%
  st_transform(4326) %>%
  st_join(access_10mw, st_intersects) %>%
  filter(!is.na(d))

tracts_access <- tracts %>%
  left_join(logsum_vars, by = "geoid") %>%
  mutate(walk_10 = ifelse(geoid %in% inside_access$geoid, TRUE, FALSE)) %>%
  write_rds("data/tracts_acces.rds")
```

Here's a map of the values 
```{r access_map}
# this interactive map is much easier to create and explore
pal <- colorQuantile("GnBu", tracts_access$access_ls_n, n = 5)
bin <- colorFactor("Dark2", tracts_access$walk_10)
leaflet(tracts_access) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(group = "access", color = ~ pal(access_ls_n),
              label = ~as.character(round(access_ls_n,2))) %>%
  # addPolygons(group = "tweets", color = ~ pal(tweets_ls),
  #             label = ~as.character(round(tweets_ls,2))) %>%
  addPolygons(group = "10-minute walk", color = ~ bin(walk_10)) %>%
  addLayersControl(baseGroups = c("access", "tweets", "10-minute walk"))
```
